### 2018-11-12 papers/journal/intro.tex
Updated introduction section: Biological part is still in early writing stage. Incomplete update.

A characteristic feature of eukaryotic cells is the presence of multiple internal membrane bound organelles. 
%
These organelles exchange molecules amongst themselves in small membrane bound packets called vesicles. 
%
The recognition of cargo molecules to be loaded onto these vesicles and the recognition of the correct target organelles for each vesicle is performed by an intricate network of molecules called the vesicle transport system []. 
%
We call the set of internal organelles, together with the molecular routes that connect them through vesicle exchange, the vesicle transport network. 
%
Although vesicle mediated traffic was discovered decades ago [17], the picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident organelles within the cell, and new vesicle routes are being discovered every year [1, 3].

>>>>

\textbf{Components of the vesicle transport network}:
The vesicle transport network is used by cells to transport molecules across compartments.
% 
The mechanism of vesicle transport can roughly be broken down into two steps: budding of vesicles from the source compartment and their fusion with the target compartment.
%
Vesicle budding involves three steps:
\begin{enumerate}
	\item  \textbf{Sorting of cargo molecules by cytosolic adaptor molecules}. 
	 Cargo adaptors bind to short linear stretches of amino acids in the cargo molecules called sorting signals and thereby sequester them for packaging. 
	 %
	 Different adaptor proteins act at different compartments, for eg., sec23/24 is required
	for packaging vesicles that form at the ER, the AP2 adaptor is used for packaging vesicles at the plasma membrane, etc. 
	%
	Adaptors are recruited to the source compartments through interactions with
	the lipids that form the source compartment's membrane, and interactions with Arf/Sar GTPases which signal the initiation of vesicle formation~\cite{paczkowski2015cargo}.

	\item \textbf{Membrane deformation, which involves cytosolic coat proteins}.
	   Coat proteins are recruited to the forming vesicle through their interaction with adaptor proteins. 
	   %
	   These are naturally curved proteins, and their assembly on membranes leads to membrane deformation.
	   %
	   A single kind of coat protein can bind to multiple kinds of adaptors, thereby allowing the packaging of multiple types of cargo within the same vesicle. 
	   %
	   Clathrin, COP2 and COP1 are the three major kinds of coat proteins
	   in eukaryotic cells. 
	   %
	   The clathrin coat operates at multiple compartments, and can interact with
	   multiple adaptors proteins, whereas the COP2 and COP1 coat are only involved in vesicle formation at a single compartment; the ER and the Golgi respectively~\cite{faini2013vesicle}

	 \item \textbf{Pinching of vesicles from the source compartment by dynamin}.
	 The final step in vesicle formation is its piching off from the source compartment.
	 %
	 This process requires ATP and is performed by the cytosolic dynamin proteins, which are recruited to the vesicle neck by the coat proteins~\cite{cocucci2014dynamin}
\end{enumerate}

Vesicle fusion involves two steps:
\begin{enumerate}
	\item \textbf{Recognition of the correct target compartment}. 
	      This is a complex process involving Rab
	      GTPases and tethering complexes:

	    \begin{enumerate}[label=(\roman*)]
	    	\item \textbf{Rab GTPases}  are small proteins which occur in two forms: the membrane associated and active
	    	GTP bound form, and the cytosolic and inactive GDP bound forms. 
	    	%
	    	Different compartments of the
	    	cell are associated with a different type of Rab protein. In their active form, Rabs recruit many
	    	downstream proteins, including tethering complexes, fusion regulators, motor proteins, sorting
	    	adaptors, etc, to the membrane they are associated with. 
	    	%
	    	The proteins that are recruited by active
	    	Rabs are called Rab effectors.
	    	%
	    	Rab activity is controlled by two proteins: GEFs turn Rabs on and
	    	GAPs turn them off. 
	    	%
	    	Each Rab has its own specific GEF and GAP. In some cases, GEFs and GAPs
	    	themselves can be Rab effectors, thus generating feed-forward and feedback loops. 
	    	%
	    	In their cytosolic, inactive form, Rabs are complexed with GDI proteins, and it is in this form that Rabs are presented to membrane bound GEFs for activation~\cite{muller2018molecular}

	    	\item \textbf{Tethering complexes} sequester vesicles to their target compartments and regulate SNARE
	    	proteins, which enable vesicle fusion. 
	    	%
	    	Tethers are believed to bind to SNAREs and function as
	    	chaperones for SNARE complex assembly which is the ultimate step in membrane fusion. 
	    	%
	    	There are two types of tethering complexes: multi subunit tethers and long coiled-coil tethers:

	    	\begin{enumerate}[label=(\alph*)]
	    		\item \textbf{Coiled coil} tethers are long, single subunit proteins. 
	    		%
	    		These tethers are believed to bind the
	    		vesicle on one end and the target compartment on the other end, thereby bridging the two before
	    		fusion.
	    		%
	    		A clear mechanism for action of these tethers is missing, but based on their interactions with
	    		other proteins of the traffic machinery, some hypotheses as to their mechanism of action have been
	    		put forth. 
	    		%
	    		Golgins, coiled coil tethers that are anchored at the Golgi by a transmembrane domain, contain a large number of binding sites for Rab GTPases. 
	    		%
	    		Thus, golgins might capture vesicles by
	    		binding vesicle-associated Rab proteins. 
	    		%
	    		Golgins also contain domains which can sense membrane
	    		curvature, which they could be using to recognize vesicles. 
	    		%
	    		They might also be interacting directly
	    		with vesicle SNAREs~\cite{baker2016chaperoning}.

	    		\item \textbf{Multi-subunit tethering complexes (MTCs)} are composed of three or more different
	    		subunits.
	    		%
	    		MTCs are known to interact with vesicle coat proteins, Rab GTPases, SNAREs and SM proteins.
	    		%
	    		For example, the HOPS complex, which is required for homotypic (fusion of two identical
	    		membranes) and heterotypic fusion in the endo-lysosomal system, was shown to tether membranes
	    		through its interactions with the membrane-associated Rab GTPase Ypt7, acidic phospholipids and
	    		SNAREs. 
	    		%
	    		The HOPS complex binds both individual SNAREs and SNARE complexes. It also
	    		seems to protect assembling trans-SNARE complexes from premature disassembly. 
	    		%
	    		Another example is given by the Dsl1 complex, which is anchored to the ER membrane through interactions
	    		with t-SNAREs. 
	    		%
	    		At the other end, the Dsl1 complex contains multiple binding sites for COPI
	    		(vesicle coat that is present on vesicles produced by the Golgi). 
	    		%
	    		This structure suggests that the Dsl1
	    		complex functions as a tether connecting COPI-coated vesicles to their target organelle, the ER~\cite{baker2016chaperoning}.

	    	\end{enumerate}

      \end{enumerate}

	  \item \textbf{Fusion of vesicle with the target compartment} is brought about by SNARE proteins. 
	  %
	  SNAREs
	  are defined by a 60- to 70-residue SNARE motif. Most SNAREs are anchored to the membrane by
	  their C-terminal transmembrane helices. 
	  %
	  The formation of a productive SNARE complex requires
	  the formation of a four-helix bundle containing four different SNARE motifs. Usually, one of these
	  motifs is contributed by the SNARE on the vesicle membrane (v-SNAREs), and the other three
	  motifs by SNAREs on the membrane of the tartget compartment (t-SNAREs). 
	  %
	  Alternatively, they
	  can be defined by their amino acid sequences as R-, Qa-, Qb-, and Qc-SNAREs. A few SNARE
	  proteins, such as SNAP25, contain both Qb- and Qc-SNARE motifs.
	  %
	  Most v-SNAREs are RSNAREs, and most t-SNAREs are Q-SNAREs~\cite{yoon2018snare}.

\end{enumerate}

Many SNAREs also have N-terminal domains that regulate SNARE complex assembly and/or
interact with other parts of the vesicle fusion machinery. 
%
Qa-SNAREs have an amino terminal
domain called the Habc domain, which can fold back on the SNARE domain and hold the QaSNARE
in an inactive state. 
%
This inactivated state is stabilized by interactions with SM proteins~\cite{yoon2018snare}.

A subset of R-SNAREs (the longin SNAREs) possess an amino-terminal domain called the longin domain. 
%
In addition to regulation of SNARE activity as in the case of the Qa-SNARE Habcdomains,
these longin domains also regulate the localization of longin SNAREs through their
interaction with adaptor proteins~\cite{daste2015structure}.

Different vesicle-target compartment pairs in the cell are associated with unique SNARE
complexes.
%
Membrane fusion converts the trans-SNARE complex (v- and t- SNAREs on opposite
membranes) into a cis-SNARE complex (both v- and t- SNAREs on the same membrane).
Disassembly of cis-SNARE complexes requires Sec17 and Sec18 (the yeast SNAP and NSF
protein, respectively). 
%
This process requires energy which is released by the hydrolysis of ATP.

\textbf{SNAREs are regulated by SM proteins}, which have three modes of action: SM proteins can hold
the Qa-SNAREs in an autoinhibited state and prevent or postpone their assembly into SNARE
complexes.
%
 Secondly, some SM proteins have been seen to act as a template upon which a halfzippered
complex between the Qa- and R-SNAREs (an early SNARE complex intermediate) can
form. 
%
In this mode, by choosing an R-SNARE located on the opposite membrane, SM proteins
could be imposing a filter inhibiting the formation of futile cis-SNARE complexes and promoting
trans-SNARE complexes.
%
Finally, SM proteins bind the four-helix bundles formed by assembled
SNARE complexes.
%
 This mode of binding of SM proteins might protect assembled SNARE
complexes from premature disassembly by NSF and SNAPs and/or might stimulate fusion directly~\cite{baker2016chaperoning}.
%
SM proteins have also been shown to interact with tethers~\cite{yoon2018snare}.

\textbf{Major paths in the vesicle transport network}:
Molecules traverse the cell in a series of vesicles, and there are two major such routes of transport in all eukaryotic cells:
\begin{enumerate}
	\item the secretory route, which takes proteins from the ER, their site of production, to the plasma
	membrane, from where they are secreted out of the cell.
	\item the endocytic route, which takes proteins from the outside of the cell through the plasma
	membrane to the endocytic compartments, where they are digested.
\end{enumerate}

Other paths are used for cross talk between the secretory and the endocytic routes.
%
There are also cyclic paths for recycle of proteins which are necessarily dislodged during the process of vesicle traffic back to their original compartments.

\textbf{Biological Model of VTS}
%
A characteristic feature of eukaryotic cells is the presence of multiple internal membrane bound organelles. 
%
These organelles exchange molecules amongst themselves in small membrane bound packets called vesicles. 
%
The recognition of cargo molecules to be loaded onto these vesicles and the recognition of the correct target organelles for each vesicle is performed by an intricate network of molecules called the vesicle transport system []. 
%
We call the set of internal organelles, together with the molecular routes that connect them through vesicle exchange, the vesicle transport network. 
%
Although vesicle mediated traffic was discovered decades ago [17], the picture of the vesicle transport network is far from complete; we do not yet know how many of the cellular proteins reach their resident organelles within the cell, and new vesicle routes are being discovered every year [1, 3].
